Generalization Errors of Hard Parameter Sharing: High-dimensional Asymptotics and Implications

### 2020-09-19 arxiv/proof_same_cov.tex
sec 3

	%\noindent\textbf{Dealing with the empirical minimizer.}
	%eet $\hat {\cal W}$ be the minimizer of $g$, and denote $\hat Q:= \hat{\cal W}^{\top} (\hat{\cal W}\hat{\cal W}^{\top})^{-1} \hat{\cal W} $.
	%We show that $\norm{\hat{\cW}^{\top} (\hat{\cW}\hat{\cW}^{\top})^{-1} \cW - U_r U_r^{\top}} \le o(1)$ w.h.p.
	%\be\label{Q-Q}\|Q_0^{-1}\hat Q - \id\|_F = \oo(1) \quad \text{w.h.p.}\ee
	%Since $\hat{\cW}$ is the minimizer of $g(\cdot)$, we have that
	%\begin{align*}
	%	g(U_r) - g(\hat{\cW}) = \bigabs{g(\hat{\cW}) - g(U_r)} &= n \cdot \bigabs{\inner{{B^{\star}}^{\top} \Sigma B^{\star}}{U_r U_r^{\top} - \hat{\cW}^{\top} (\hat{\cW}\hat{\cW}^{\top})^{-1} \hat{\cW}}}\\
	%	&\ge n \cdot \sigma_{\min}({B^{\star}}^{\top} \Sigma B^{\star}) \cdot \bignormFro{U_r U_r^{\top} - \hat{\cW}^{\top} (\hat{\cW}\hat{\cW}^{\top})^{-1} \hat{\cW}}.
	%\end{align*}
	Combining equation \eqref{eq_g_gap}, \eqref{eq_g_gap1}, and \eqref{eq_g_gap2}, we conclude that
	\begin{align*}
		\bignormFro{U_r U_r^{\top} - U_{\hat{A}} U_{\hat{A}}^{\top}}
		\le \frac{p^{-c} \rho \normFro{\Sigma^{1/2} B^{\star}}^2 + p^{-1/2 + \e} \cdot \sigma^2 \rho t}{\lambda_{\min}^2(\Sigma^{1/2} B^{\star}) ((\sqrt{\rho} - 1)^2 - \rho \cdot p^{-c}) + \sigma^2 (1 - p^{-1/2 + \e})}.
	\end{align*}
	Hence for large enough $p$, we obtain that Claim \ref{claim_opt_dist} holds.
	The proof is complete.
	%The second step uses the fact that $U_r U_r^{\top}$ is the global minimum of $\ex{g(\cdot)}$.
	%In fact, if \eqref{Q-Q} does not hold, then using the condition $\lambda_{\min}((B^{\star})^\top\Sigma B^{\star})\gtrsim \sigma^2$ and that $\delta_{\vari}(\cal W)=\OO(\sigma^2)$ by  Lemma \ref{lem_minv}, we obtain that
	%$$   \val(\hat {Q}) + t  \sigma^2 > (\val( Q_0) + t \sigma^2 )\cdot (1+\oo(1)) \ \Rightarrow \ g( \hat{\cal W})>g( {\cal W}_0),$$
	%where $\cal W_0\in \R^{r\times t}$ is a matrix such that  $ \cal W_0^{\top} (\cal W_0\cal W_0^{\top})^{-1} \cal W_0=Q_0$. Hence $\hat {\cal W}$ is not a minimizer, which leads to a contradiction.
	%with high probability
	%\begin{align}
	%g(\cal W)&= n \bignorm{\Sigma^{1/2} B^{\star} (\cW^{\top} (\cW \cW^{\top})^{-1} \cW - \id_{t\times t})}^2 \cdot (1+\oo(1)) + \sigma^2 (n\cdot t - p \cdot r)+\OO(\sigma^2 p^{1/2+\e}) \nonumber \\
	%&=\ex{g(\cW)} \cdot (1+\oo(1)) + \OO(\sigma^2 p^{1/2+\e}). \label{eq_est_g}
	%\end{align}

>>>>

	%\noindent\textbf{Dealing with the empirical minimizer.}
	%eet $\hat {\cal W}$ be the minimizer of $g$, and denote $\hat Q:= \hat{\cal W}^{\top} (\hat{\cal W}\hat{\cal W}^{\top})^{-1} \hat{\cal W} $.
	%We show that $\norm{\hat{\cW}^{\top} (\hat{\cW}\hat{\cW}^{\top})^{-1} \cW - U_r U_r^{\top}} \le o(1)$ w.h.p.
	%\be\label{Q-Q}\|Q_0^{-1}\hat Q - \id\|_F = \oo(1) \quad \text{w.h.p.}\ee
	%Since $\hat{\cW}$ is the minimizer of $g(\cdot)$, we have that
	%\begin{align*}
	%	g(U_r) - g(\hat{\cW}) = \bigabs{g(\hat{\cW}) - g(U_r)} &= n \cdot \bigabs{\inner{{B^{\star}}^{\top} \Sigma B^{\star}}{U_r U_r^{\top} - \hat{\cW}^{\top} (\hat{\cW}\hat{\cW}^{\top})^{-1} \hat{\cW}}}\\
	%	&\ge n \cdot \sigma_{\min}({B^{\star}}^{\top} \Sigma B^{\star}) \cdot \bignormFro{U_r U_r^{\top} - \hat{\cW}^{\top} (\hat{\cW}\hat{\cW}^{\top})^{-1} \hat{\cW}}.
	%\end{align*}
	Combining equation \eqref{eq_g_gap}, \eqref{eq_g_gap1}, and \eqref{eq_g_gap2}, we conclude that
	\begin{align*}
		\bignormFro{U_r U_r^{\top} - U_{\hat{A}} U_{\hat{A}}^{\top}}
		\le \frac{p^{-c} \rho \normFro{\Sigma^{1/2} B^{\star}}^2 + p^{-1/2 + \e} \cdot \sigma^2 \rho t}{\lambda_{\min}^2(\Sigma^{1/2} B^{\star}) ((\sqrt{\rho} - 1)^2 - \rho \cdot p^{-c}) + \sigma^2 (1 - p^{-1/2 + \e})}.
	\end{align*}
	Hence for large enough $p$, we obtain that Claim \ref{claim_opt_dist} holds.
	The proof is complete.

	\paragraph{Proof of Claim \ref{claim_pred_err}.}

	\paragraph{Proof of Theorem \ref{thm_many_tasks}.}
	%In sum, we have solved that $\hat{\beta}_i^{\MTL}=B^{\star}\left( U_r v_i +\oo(1)\right)$.
	Now we are ready to finish the proof.
	Similar to the proof of equation \eqref{eq_est_g1} and \eqref{eq_est_g2}, we have that
	\begin{align}
		L(\hat{\beta}_t^{\MTL}) = \exarg{\varepsilon_1, \dots, \varepsilon_t}{L(\hat{\beta}_t^{\MTL})} (1 + O(p^{-1/2+c})) + O(\sigma^2 \cdot p^{-1/2 + c}). \label{eq_multiple_last1}
	\end{align}
	Using equation \eqref{eq_ex_pred}, we have that
	\begin{align}
		\exarg{\varepsilon_1, \dots, \varepsilon_t}{\hat{\beta}_t^{\MTL}} &= \bignorm{\Sigma^{1/2} \left((B^\star \hat{\cal W}^\top) (\hat{\cal W}\hat{\cal W}^{\top})^{-1} \hat W_t - \beta_t \right) }^2
		+ \sigma^2  \hat W_t^{\top} (\hat{\cal W}\hat{\cal W}^{\top})^{-1} \hat W_t \cdot \bigtr{\Sigma (X^{\top}X)^{-1}} \nonumber \\
		&= \bignorm{\Sigma^{1/2} \bigbrace{B^{\star} U_r v_t-\beta_t}}^2 + \oo\left(\bigtr{{B^{\star}}^{\top} \Sigma B^{\star}}\right) + \sigma^2(\norm{v_t}^2 + o(1)) \bigtr{\Sigma (X^{\top}X)^{-1}} \cdot (1+\oo(1)) \nonumber \\
		&= \bignorm{\Sigma^{1/2} \bigbrace{B^{\star} U_r v_t-\beta_t}}^2 + \frac{\sigma^2}{\rho-1}\norm{v_t}^2 + \oo \left( \bigtr{{B^{\star}}^{\top} \Sigma B^{\star}} + \sigma^2\right), \label{eq_multiple_last2}
	\end{align}
	In the second step, we use that $\norm{\hat{\cW}^{\top} (\cW \cW^{\top})^{-1} \cW - U_r U_r^{\top}} \le o(1)$ w.h.p., which implies that
	\[ \bignorm{U_r v_t - \hat{\cW}^{\top} (\cW \cW^{\top})^{-1} \hat{\cW}_t} \le o(1) \quad \text{ and } \quad \bigabs{\hat{\cW}_t^{\top} (\hat{\cW} \hat{\cW}^{\top})^{-1} \hat{W}_t - \norm{v_t}^2} \le o(1). \]
  In the last step, we use Lemma \ref{lem_minv}, that is, $\bigtr{\Sigma (X^{\top} X)^{-1}} = \sigma^2 \cdot (1 + o(1))/ (\rho - 1)$.
	Combining equation \eqref{eq_multiple_last1}, \eqref{eq_multiple_last2}, and $\te(\hat{\beta}_t^{\STL})=\frac{\sigma^2}{\rho-1} \cdot \left( 1+\oo(1)\right)$, we conclude the proof.
	%The second step uses the fact that $U_r U_r^{\top}$ is the global minimum of $\ex{g(\cdot)}$.
	%In fact, if \eqref{Q-Q} does not hold, then using the condition $\lambda_{\min}((B^{\star})^\top\Sigma B^{\star})\gtrsim \sigma^2$ and that $\delta_{\vari}(\cal W)=\OO(\sigma^2)$ by  Lemma \ref{lem_minv}, we obtain that
	%$$   \val(\hat {Q}) + t  \sigma^2 > (\val( Q_0) + t \sigma^2 )\cdot (1+\oo(1)) \ \Rightarrow \ g( \hat{\cal W})>g( {\cal W}_0),$$
	%where $\cal W_0\in \R^{r\times t}$ is a matrix such that  $ \cal W_0^{\top} (\cal W_0\cal W_0^{\top})^{-1} \cal W_0=Q_0$. Hence $\hat {\cal W}$ is not a minimizer, which leads to a contradiction.
	%with high probability
	%\begin{align}
	%g(\cal W)&= n \bignorm{\Sigma^{1/2} B^{\star} (\cW^{\top} (\cW \cW^{\top})^{-1} \cW - \id_{t\times t})}^2 \cdot (1+\oo(1)) + \sigma^2 (n\cdot t - p \cdot r)+\OO(\sigma^2 p^{1/2+\e}) \nonumber \\
	%&=\ex{g(\cW)} \cdot (1+\oo(1)) + \OO(\sigma^2 p^{1/2+\e}). \label{eq_est_g}
	%\end{align}
